---
title: "Workflow with Caret"
author: "pauke"
date: "1/26/2019"
output: html_document
---

<style type="text/css">

body{ /* Normal  */
      font-size: 21px;
  }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(DiagrammeR)
```

## Foreword

This is a workflow about [caret](https://topepo.github.io/caret/), which is a creating predicted models collecting toolkit framework, integrating all activities related to model development from other R packages. Caret can be also treated as a tool to get started or get familiar with machine learning, especially in supervised and unsupervised learning part that is good at. 

The article is about the usual thought of using caret and machine learning, more focusing on general use situation. And I'll try to offer a holistic view for machine learning beginners and the way to find more concrete field. That means I would not present the specific algorithms, tuning methods or fitted fields.

## Structure

All tools in caret can be classified in the [ducument](https://topepo.github.io/caret/) like this:

> - data splitting
> - pre-processing
> - feature selection
> - model tuning using resampling
> - variable importance estimation

Link to the crutial function in each part, all the part of caret workflow is:


```{r echo=FALSE, message=FALSE, warning=FALSE}

DiagrammeR::mermaid('
graph LR
D_E[Data Exploration/Visualization]
P[Preprocessing]
D_S[Data Splitting]
M[Modeling]
M_E[Model Evaluation]

D_Es("featureplot()<br>dotPlot()<br>lift()<br>plotClassProbs()")
Ps("nearZeroVar()<br>findCorrelation()<br>findLinearCombos()<br>preprocess()")
D_Ss("createDataPartition()<br>createFolds()")
Ms("train()<br>trainControl()")
M_Es("confusionMatrix()<br>resample()<br>postResample()")

D_E --> P 
P --> D_S
D_S --> M
M --> M_E

D_E --- D_Es
P --- Ps
D_S --- D_Ss
M --- Ms
M_E --- M_Es

classDef default fill:#ffffe6,stroke:#333,stroke-width:2px

style D_E stroke-width:4px
style P stroke-width:4px
style D_S stroke-width:4px
style M stroke-width:4px
style M_E stroke-width:4px

', height =500, width =950
)

```


By the way,as I said in foreword, this post would like more pay attention on the process of caret using. The part of data exploration is deeply dependent on the familiarity about the data you get to deal with, and can be replaced by another professional packages. If you are more interesting on the visualization tools of caret, just jump to this [chapter](http://topepo.github.io/caret/visualizations.html).




## Preprocessing

This part is based on the conclution of your data exploration. You need to consider what situation of your data that you need to deal with before starting build your model. According to the situation of the function facing to, preprocessing function of caret can be sorted as:


category|function
----------|--------
categorical/dichotomous variables| dummyVars()
unbalance| nearZeroVar()










